---
title: 'Geog6300: Final Project'
author: "Jerry Shannon"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
```

## Introduction

For this final project, you will use climatological and demographic data on census tracts in the Atlanta metropolitan area to answer a research question of your choosing. You'll do so using techniques for descriptive and inferential analysis that we've worked on throughout this course. 

## The dataset

The dataset for your final includes a lot of variables about tracts in the Atlanta area. You can see a full list on the README page in this repo. The climatological data is taken from Daymet and the demographic information is from the U.S. Census American Community Survey. There's also classifications of racial composition from the [MixedMetro project](http://mixedmetro.com).  

```{r message=FALSE}
atl_data<-st_read("data/atl_data.gpkg",layer="acsdata_atl_climate", quiet=TRUE) %>%
  rename(geometry=geom)
```

## Research question and variables (5 points)
_Your first task is to develop a single research question that relies on 3-5 variables from this dataset. This question will probably be quite narrow, and you can only use the data provided to you. I prefer that you take a "less is more" approach here--focus on asking a question you know how to answer. In the space below, state your question and the reasons you think it is of interest._


## Variables and methods (10 points)
_Describe the variables you'll use for your analysis and your reasons for including them. Then describe the analysis you'll be doing with these variables and explain how these helps answer the research question. These should include only methods we've covered in class (unless you've gotten explicit persmission to use others). Remember to use diagnostics such as analysis of model residuals or post-hoc testing if appropriate._


## Descriptive analysis (10 points)
_Describe the central tendency (e.g., mean, median, mode), statistical distribution (e.g., standard deviation and/or IQR), and spatial distribution of each variable using the techniques we have covered in class. Also assess the normality of variables if appropriate. Describe any notable patterns you see in these variables based on this analysis._

## Inferential statistics (10 points)
_Use at least one of the inferential statistical tests we've covered in class to analyze the data. You should also interpret the results of any tests you run. For example, for a regression, what is the magnitude, direction, and significance of your coefficients, and how would you explain what they tell you in plain language?_

## Conclusion (5 points)
_Based on your analysis, what answer can you give to your research question? Make sure to base your conclusions of the results of your analysis._